---
title: "Final Project 3rd Data Visual (Glit)"
author: "Eliot Ozaki, Glit Hanpanitkitkan, Ethan Kawahara, Andrew Sousa"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(countrycode)
```


```{r data}
#setwd(dirname(rstudioapi::getActiveDocumentContext()$path)) # Requires RStudio

gdp_data = read.csv("Datasets/GDP.PCAP.csv")
demog_data = read.csv("Datasets/UNdata_secondary_enrollment.csv")
gdpmeta_data = read.csv("Datasets/Metadata_Country_API_NY.GDP.PCAP.CD_DS2_en_csv_v2_9803.csv")
```

# Final Project


### Introductory Analysis
```{r}
# Selecting only X2000-X2005 from the GDP data
gdp_data = gdp_data %>% select(1:2, 44:50)
# Renaming the columns
colnames(gdp_data) = c("CountryName", "CountryCode", "1999","2000", "2001", "2002", "2003", "2004", "2005")
print(sum(is.na(gdp_data)))

# Printing num.na in each column:
print(colSums(is.na(gdp_data)))
head(gdp_data)
```


```{r}
tail(demog_data, 15)
## Dropping the last 6 rows
demog_data = demog_data[1:2338,]

## Changing values of '-' to 0
demog_data = demog_data %>% mutate_all(funs(ifelse(. == "-", 0, .)))

demog_data$Value = as.numeric(demog_data$Value)
print(colSums(is.na(demog_data)))

print(summary(demog_data))

demog_data = demog_data %>% select(-Value.Footnotes)

# spreading demog data by Year
demog_data = demog_data %>% spread(key = Year, value = Value)
print(sum(is.na(demog_data)))
head(demog_data)


for (i in 1:6) {
  demog_data$`1999` = ifelse(is.na(demog_data$`1999`), demog_data$`2000`, demog_data$`1999`)
  demog_data$`2000` = ifelse(is.na(demog_data$`2000`), demog_data$`2001`, demog_data$`2000`)
  demog_data$`2001` = ifelse(is.na(demog_data$`2001`), demog_data$`2002`, demog_data$`2001`)
  demog_data$`2002` = ifelse(is.na(demog_data$`2002`), demog_data$`2003`, demog_data$`2002`)
  demog_data$`2003` = ifelse(is.na(demog_data$`2003`), demog_data$`2004`, demog_data$`2003`)
  demog_data$`2004` = ifelse(is.na(demog_data$`2004`), demog_data$`2005`, demog_data$`2004`)

  demog_data$`2005` = ifelse(is.na(demog_data$`2005`), demog_data$`2004`, demog_data$`2005`)
  demog_data$`2004` = ifelse(is.na(demog_data$`2004`), demog_data$`2003`, demog_data$`2004`)
  demog_data$`2003` = ifelse(is.na(demog_data$`2003`), demog_data$`2002`, demog_data$`2003`)
  demog_data$`2002` = ifelse(is.na(demog_data$`2002`), demog_data$`2001`, demog_data$`2002`)
  demog_data$`2001` = ifelse(is.na(demog_data$`2001`), demog_data$`2000`, demog_data$`2001`)
  demog_data$`2000` = ifelse(is.na(demog_data$`2000`), demog_data$`1999`, demog_data$`2000`)

}


## Printing cols where all year values are NA
print(sum(is.na(demog_data)))
demog_data = demog_data %>% select(-Unit)

```


```{r}
print(summary(gdpmeta_data))

print(colSums(is.na(gdpmeta_data)))

gdpmeta_data$Region = ifelse(gdpmeta_data$Region == "", "Other", gdpmeta_data$Region)
gdpmeta_data$IncomeGroup = ifelse(gdpmeta_data$IncomeGroup == "", "Other", gdpmeta_data$IncomeGroup)

## Renaming Country Code
colnames(gdpmeta_data)[1] = "CountryCode"
```

```{r}
## Merging gdp and gdpmeta datasets
gdp_data = merge(gdp_data, gdpmeta_data, by = "CountryCode")

summary(gdp_data)
```

##### Our Question
Our question will be: How did income level affect ratio of women to men who enroll in secondary education from 1999-2005? We will answer this question using GDP and demographic data, looking at the general change, change by region, change by income group of a country, and change by large country. The hypothesis is that higher income countries will have a higher ratio of women to men who enroll in secondary education. The data we will be looking at is from the world bank (income) and UN data (enrollment in secondary education).

 
```{r}
demog_data_long <- demog_data %>% 
  pivot_longer(cols = c(4:10), names_to = "Year", values_to = "Enrollment")

demog_data_long$iso3c <- countrycode(demog_data_long$Country.or.Area, origin = "country.name", destination = "iso3c")

demog_data_long <- demog_data_long %>% 
  inner_join(gdp_data, by = c("iso3c" = "CountryCode")) %>% 
  rename("CountryCode" = iso3c) %>% 
  relocate(CountryCode, .before = 1)

demog_data_long <- demog_data_long %>% 
  mutate(Income_Low_High = case_when(IncomeGroup %in% c("Low income", "Lower middle income") ~ "Lower income",
                                     IncomeGroup %in% c("High income", "Upper middle income") ~ "Upper income",
                                     TRUE ~ NA))


```
```{r}

demog_data_long %>% 
  group_by(Subgroup, Year, Income_Low_High) %>% 
  summarize(Average_Enrollment = mean(Enrollment, na.rm = TRUE)) %>% 
  mutate(Group = paste(Subgroup, Income_Low_High, sep = " ")) %>%  # Combine variables into a single grouping
  filter(Income_Low_High != "Other") %>% 
  ggplot(aes(x = Year, y = Average_Enrollment, fill = Group)) +
  geom_col(position = position_dodge2(padding = 0.2), alpha = 0.7) +
  geom_line(aes(color = Group, group = Group), size = 0.7, show.legend = FALSE) + 
  labs(title = "How does a country's income level affect the enrollment rates of \nmen and women between from 1999 to 2005?",
       subtitle = "Average global enrollment over time by gender and income level",
       x = "Year",
       y = "Average Enrollment",
       fill = "Gender and Income Level",
       caption = "Source: UN Data enrollment in secondary education") +
  scale_y_continuous(
    expand = c(0, 0),  # Removes extra space at the bottom
    limits = c(0, 1600000),  # Manually set upper limit to a higher value (adjust this value if needed)
    breaks = seq(0, 2000000, by = 200000)  # Set the breaks according to your desired range
  ) +  
  scale_fill_manual(values = c("Female Lower income" = "lightpink", 
                               "Female Upper income" = "magenta", 
                               "Male Lower income" = "lightblue", 
                               "Male Upper income" = "blue")) + 
  scale_color_manual(values = c("Female Lower income" = "lightpink", 
                               "Female Upper income" = "magenta", 
                               "Male Lower income" = "lightblue", 
                               "Male Upper income" = "blue")) +
  theme_bw() +
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))  # Adjust margins to minimize space around the plot



```