---
title: "FinalProject"
author: "Eliot Ozaki, Glit Hanpanitkitkan, Ethan Kawahara, Andrew Sousa"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs}
library(tidyverse)
library(ggplot2)
library(dplyr)
```


```{r data}
gdp_data = read.csv("Datasets/GDP.PCAP.csv")
demog_data = read.csv("Datasets/UNdata_secondary_enrollment.csv")
gdpmeta_data = read.csv("Datasets/Metadata_Country_API_NY.GDP.PCAP.CD_DS2_en_csv_v2_9803.csv")
```

# Final Project


### Introductory Analysis
```{r}
# Selecting only X2000-X2005 from the GDP data
gdp_data = gdp_data %>% select(1:2, 44:50)
# Renaming the columns
colnames(gdp_data) = c("CountryName", "CountryCode", "1999","2000", "2001", "2002", "2003", "2004", "2005")
print(sum(is.na(gdp_data)))

# Printing num.na in each column:
print(colSums(is.na(gdp_data)))
head(gdp_data)
```


```{r}
tail(demog_data, 15)
## Dropping the last 6 rows
demog_data = demog_data[1:2338,]

## Changing values of '-' to 0
demog_data = demog_data %>% mutate_all(funs(ifelse(. == "-", 0, .)))

demog_data$Value = as.numeric(demog_data$Value)
print(colSums(is.na(demog_data)))

print(summary(demog_data))

demog_data = demog_data %>% select(-Value.Footnotes)

# spreading demog data by Year
demog_data = demog_data %>% spread(key = Year, value = Value)
print(sum(is.na(demog_data)))
head(demog_data)


for (i in 1:6) {
  demog_data$`1999` = ifelse(is.na(demog_data$`1999`), demog_data$`2000`, demog_data$`1999`)
  demog_data$`2000` = ifelse(is.na(demog_data$`2000`), demog_data$`2001`, demog_data$`2000`)
  demog_data$`2001` = ifelse(is.na(demog_data$`2001`), demog_data$`2002`, demog_data$`2001`)
  demog_data$`2002` = ifelse(is.na(demog_data$`2002`), demog_data$`2003`, demog_data$`2002`)
  demog_data$`2003` = ifelse(is.na(demog_data$`2003`), demog_data$`2004`, demog_data$`2003`)
  demog_data$`2004` = ifelse(is.na(demog_data$`2004`), demog_data$`2005`, demog_data$`2004`)

  demog_data$`2005` = ifelse(is.na(demog_data$`2005`), demog_data$`2004`, demog_data$`2005`)
  demog_data$`2004` = ifelse(is.na(demog_data$`2004`), demog_data$`2003`, demog_data$`2004`)
  demog_data$`2003` = ifelse(is.na(demog_data$`2003`), demog_data$`2002`, demog_data$`2003`)
  demog_data$`2002` = ifelse(is.na(demog_data$`2002`), demog_data$`2001`, demog_data$`2002`)
  demog_data$`2001` = ifelse(is.na(demog_data$`2001`), demog_data$`2000`, demog_data$`2001`)
  demog_data$`2000` = ifelse(is.na(demog_data$`2000`), demog_data$`1999`, demog_data$`2000`)

}


## Printing cols where all year values are NA
print(sum(is.na(demog_data)))
demog_data = demog_data %>% select(-Unit)

```


```{r}
print(summary(gdpmeta_data))

print(colSums(is.na(gdpmeta_data)))

gdpmeta_data$Region = ifelse(gdpmeta_data$Region == "", "Other", gdpmeta_data$Region)
gdpmeta_data$IncomeGroup = ifelse(gdpmeta_data$IncomeGroup == "", "Other", gdpmeta_data$IncomeGroup)

## Renaming Country Code
colnames(gdpmeta_data)[1] = "CountryCode"
```

```{r}
## Merging gdp and gdpmeta datasets
gdp_data = merge(gdp_data, gdpmeta_data, by = "CountryCode")

summary(gdp_data)
```

##### Our Question
Our question will be: How did income level affect ratio of women to men who enroll in secondary education from 1999-2005? We will answer this question using GDP and demographic data, looking at the general change, change by region, change by income group of a country, and change by large country. The hypothesis is that higher income countries will have a higher ratio of women to men who enroll in secondary education. The data we will be looking at is from the world bank (income) and UN data (enrollment in secondary education).

 


