---
title: "Final Project Data Visuals (Ethan)"
author: "Eliot Ozaki, Glit Hanpanitkitkan, Ethan Kawahara, Andrew Sousa"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(readr)
library(fastmap)
library(shiny)
library(htmltools)
library(countrycode)
```



```{r data}


gdp_data = read.csv("Datasets/GDP.PCAP.csv", skip = 3)
gdp_data2 = read.csv("Datasets/GDP.PCAP.csv", skip = 3)
demog_data = read.csv("Datasets/UNdata_secondary_enrollment.csv")
demog_data2 = read.csv("Datasets/UNdata_secondary_enrollment.csv")
gdpmeta_data = read.csv("Datasets/Metadata_Country_API_NY.GDP.PCAP.CD_DS2_en_csv_v2_9803.csv")
gdpmeta_data2 = read.csv("Datasets/Metadata_Country_API_NY.GDP.PCAP.CD_DS2_en_csv_v2_9803.csv")
```

# Final Project


### Introductory Analysis
```{r}
# Selecting only X2000-X2005 from the GDP data
gdp_data = gdp_data %>% select(1:2, 44:50)
# Renaming the columns
colnames(gdp_data) = c("CountryName", "CountryCode", "1999","2000", "2001", "2002", "2003", "2004", "2005")
print(sum(is.na(gdp_data)))

# Printing num.na in each column:
print(colSums(is.na(gdp_data)))
head(gdp_data)

for (i in 1:6) {
  gdp_data$`2005` = ifelse(is.na(gdp_data$`2005`), gdp_data$`2004`, gdp_data$`2005`)
  gdp_data$`2004` = ifelse(is.na(gdp_data$`2004`), gdp_data$`2003`, gdp_data$`2004`)
  gdp_data$`2003` = ifelse(is.na(gdp_data$`2003`), gdp_data$`2002`, gdp_data$`2003`)
  gdp_data$`2002` = ifelse(is.na(gdp_data$`2002`), gdp_data$`2001`, gdp_data$`2002`)
  gdp_data$`2001` = ifelse(is.na(gdp_data$`2001`), gdp_data$`2000`, gdp_data$`2001`)
  gdp_data$`2000` = ifelse(is.na(gdp_data$`2000`), gdp_data$`1999`, gdp_data$`2000`)
  
  gdp_data$`1999` = ifelse(is.na(gdp_data$`1999`), gdp_data$`2000`, gdp_data$`1999`)
  gdp_data$`2000` = ifelse(is.na(gdp_data$`2000`), gdp_data$`2001`, gdp_data$`2000`)
  gdp_data$`2001` = ifelse(is.na(gdp_data$`2001`), gdp_data$`2002`, gdp_data$`2001`)
  gdp_data$`2002` = ifelse(is.na(gdp_data$`2002`), gdp_data$`2003`, gdp_data$`2002`)
  gdp_data$`2003` = ifelse(is.na(gdp_data$`2003`), gdp_data$`2004`, gdp_data$`2003`)
  gdp_data$`2004` = ifelse(is.na(gdp_data$`2004`), gdp_data$`2005`, gdp_data$`2004`)
}

for (col in 3:9) {
  if (any(is.na(gdp_data[,col]))) {
    gdp_data[,col] = ifelse(is.na(gdp_data[,col]), mean(gdp_data[,col], na.rm = TRUE), gdp_data[,col])
  }
}
```


```{r}
tail(demog_data, 15)
## Dropping the last 6 rows
demog_data = demog_data[1:2338,]

## Changing values of '-' to 0
demog_data = demog_data %>% mutate_all(funs(ifelse(. == "-", 0, .)))

demog_data$Value = as.numeric(demog_data$Value)
print(colSums(is.na(demog_data)))

print(summary(demog_data))

demog_data = demog_data %>% select(-Value.Footnotes)

# spreading demog data by Year
demog_data = demog_data %>% spread(key = Year, value = Value)
print(sum(is.na(demog_data)))
head(demog_data)


for (i in 1:6) {
  demog_data$`1999` = ifelse(is.na(demog_data$`1999`), demog_data$`2000`, demog_data$`1999`)
  demog_data$`2000` = ifelse(is.na(demog_data$`2000`), demog_data$`2001`, demog_data$`2000`)
  demog_data$`2001` = ifelse(is.na(demog_data$`2001`), demog_data$`2002`, demog_data$`2001`)
  demog_data$`2002` = ifelse(is.na(demog_data$`2002`), demog_data$`2003`, demog_data$`2002`)
  demog_data$`2003` = ifelse(is.na(demog_data$`2003`), demog_data$`2004`, demog_data$`2003`)
  demog_data$`2004` = ifelse(is.na(demog_data$`2004`), demog_data$`2005`, demog_data$`2004`)

  demog_data$`2005` = ifelse(is.na(demog_data$`2005`), demog_data$`2004`, demog_data$`2005`)
  demog_data$`2004` = ifelse(is.na(demog_data$`2004`), demog_data$`2003`, demog_data$`2004`)
  demog_data$`2003` = ifelse(is.na(demog_data$`2003`), demog_data$`2002`, demog_data$`2003`)
  demog_data$`2002` = ifelse(is.na(demog_data$`2002`), demog_data$`2001`, demog_data$`2002`)
  demog_data$`2001` = ifelse(is.na(demog_data$`2001`), demog_data$`2000`, demog_data$`2001`)
  demog_data$`2000` = ifelse(is.na(demog_data$`2000`), demog_data$`1999`, demog_data$`2000`)

}


## Printing cols where all year values are NA
print(sum(is.na(demog_data)))
demog_data = demog_data %>% select(-Unit)

```


```{r}
print(summary(gdpmeta_data))

print(colSums(is.na(gdpmeta_data)))

gdpmeta_data$Region = ifelse(gdpmeta_data$Region == "", "Other", gdpmeta_data$Region)
gdpmeta_data$IncomeGroup = ifelse(gdpmeta_data$IncomeGroup == "", "Other", gdpmeta_data$IncomeGroup)

## Renaming Country Code
colnames(gdpmeta_data)[1] = "CountryCode"
```

```{r}
## Merging gdp and gdpmeta datasets
gdp_data = merge(gdp_data, gdpmeta_data, by = "CountryCode")

#write.csv(gdp_data, "Datasets/gdp_data.csv", row.names = FALSE)


summary(gdp_data)
```

##### Our Question
Our question will be: How did income level affect ratio of women to men who enroll in secondary education from 1999-2005? We will answer this question using GDP and demographic data, looking at the general change, change by region, change by income group of a country, and change by large country. The hypothesis is that higher income countries will have a higher ratio of women to men who enroll in secondary education. The data we will be looking at is from the world bank (income) and UN data (enrollment in secondary education).


```{r}
gdp_long <- gdp_data %>%
  pivot_longer(cols = starts_with("19") | starts_with("20"), 
               names_to = "Year", 
               values_to = "GDP") %>%
  filter(!is.na(Region) & Region != "World")

# Convert 'Year' to numeric
gdp_long$Year <- as.numeric(gdp_long$Year)

# Calculate the average GDP for each region by year
average_gdp_by_region_year <- gdp_long %>%
  group_by(Region, Year) %>%
  summarize(Average_GDP = mean(GDP, na.rm = TRUE))

# Create the line plot
gdp <- ggplot(average_gdp_by_region_year, aes(x = Year, y = Average_GDP, color = Region)) +
  geom_line() +
  geom_point() +
  labs(title = "Average GDP Per Capita by Region Over Time", x = "Year", y = "Average GDP") 
```
```{r}
library(scales)  # For formatting y-axis

demog_long <- demog_data %>%
  pivot_longer(cols = starts_with("19") | starts_with("20"), 
               names_to = "Year", 
               values_to = "Value") %>%
  filter(!is.na(Value))  # Remove rows with NA values if needed

# Merge with region data from gdp_data
demog_data_with_region <- merge(demog_long, gdp_data[, c("CountryName", "Region")], 
                                by.x = "Country.or.Area", by.y = "CountryName", all.x = TRUE)

# Convert 'Year' to numeric
demog_data_with_region$Year <- as.numeric(demog_data_with_region$Year)

#write.csv(demog_data_with_region, "Datasets/demog_data_with_region.csv", row.names = FALSE)

# Calculate mean value per region and year
demog_mean_values <- demog_data_with_region %>%
  group_by(Region, Year) %>%
  summarize(Mean_Value = mean(Value, na.rm = TRUE), .groups = 'drop') %>%
  arrange(Region, Year)  # Sort by region and year

# Calculate growth relative to the first year in each region
demog_growth <- demog_mean_values %>%
  group_by(Region) %>%
  mutate(Growth = ((Mean_Value / first(Mean_Value)) * 100) - 100)  # Standardize to the first year

# Plot the growth
growth <- ggplot(demog_growth, aes(x = Year, y = Growth, color = Region)) +
  geom_line() +
  geom_point() +
  labs(title = "Growth in Secondary Education Enrollment by Region", x = "Year", y = "Enrollment Growth (%)") 
```

```{r}
#Initial layouts of RShiny app
#Let people choose which plot they want to see

ui <- fluidPage(
  titlePanel("Choose a Plot"),
  sidebarLayout(
    sidebarPanel(
      selectInput(
        "plotChoice",
        "Choose a Plot",
        choices = c("GDP" = "gdp", "Growth" = "growth")
      )
    ),
    mainPanel(
      plotOutput("chosenPlot")
    )
  )
)

server <- function(input, output){
  output$chosenPlot <- renderPlot({
    if(input$plotChoice == "gdp"){
      gdp
    } else if(input$plotChoice == "growth"){
      growth
    }
  })

}

shinyApp(ui, server)
```

